### 2011-11-18 paper.tex
TXT Little abstract for the paper
Still not sure whether this is going, but it might be a good idea to
have a write-up.

\documentclass{article}

\title{Jug: Software for Reproducible Computation}
\author{Luis Pedro Coelho}

\begin{document}
\section*{Abstract}
As computational pipelines become a bigger part of science, it is important to
ensure that the results are reproducible. All developed software should be able
to be run automatically without any user intervention.

In addition to the value to the community of being able to reproduce an
analysis, reproducible research practices allow for better control over the
project. For example, if necessary parameters to run a pipeline are kept
separately from the code that implements it (perhaps even in the researcher's
mind), this leads to error-prone analysis and opens up the possibility that
when the results are to be writtne up for publication, the researcher will no
longer be able to even completely describe the process that led to them.

For large projects, the use of multiple processors (either in the same machine
or distributed across a cluster) is necessary to obtain results in a useful
timeframe. Furthermore, it is often the case that, as the project evolves, it
becomes necessary to save intermediate results while down-stream analyses are
designed (or redesigned) and implemented. Therefore, having a single point of
entry (often in the form of a single script or a single main programme) for the
computation becomes increasingly difficult.

Jug is a software framework which solves all of these problems in a
simple way. Jug supports caching of intermediate results, distribution of
computation as tasks across a network.

Jug is written in pure Python, is completely cross-platform, and available as
free software under the MIT license. Jug is available from
\url{http://github.com/luispedro/jug}.

\section{Introduction}
\section{Methods}
\section{Example}
\section{Discussion}
\end{document}

### 2011-11-21 paper.tex
TXT Copy over section from thesis
This is a good start, but it is too succint. Also, in the thesis this
was just a section in a larger chapter. Here it needs to be broken up.
Also, it needs to be integrated into the intellectual framework of the
paper instead of just as an appendix to something unrelated.

@@
 \section{Introduction}
+Jug is a task-based framework for Python, which supports saving and sharing of
+intermediate results and parallelisation on computer clusters (or multi-core
+machines).
+
+\subsection{Task Based Architecture}
+
+Jug is designed around tasks. A task is defined as a Python function and a set
+of arguments, which may be Python values or the output of other tasks. A task
+should be a pure function, but there is no way of mandating this in the
+language.
+
+\begin{python}
+from jug import Task
+
+def count(imname):
+    ...
+    return value
+
+def mean(args):
+    return sum(args)/float(len(args))
+
+images = glob('*.png')
+counts = [Task(count, im) for im in images]
+final = Task(mean, counts)
+\end{python}
+
+\begin{figure}
+\begin{center}
+\begin{tikzpicture}
+\node (mean) at (0,2) [draw] {\code{mean}};
+\node (count0) at (5,3) [draw] {\code{count('0.png')}};
+\node (count1) at (5,2) [draw] {\code{count('1.png')}};
+\node (count2) at (5,1) [draw] {\code{count('2.png')}};
+\node (countn) at (5,0) {\ldots};
+
+\draw[->] (mean.east) -- (count0.west);
+\draw[->] (mean.east) -- (count1.west);
+\draw[->] (mean.east) -- (count2.west);
+\draw[->] (mean.east) -- (countn.west);
+\end{tikzpicture}
+
+\end{center}
+\caption{Simple Dependency Structure for Example in the Text. This assumes that
+the directory had a collection of images names 0.png, 1.png,\ldots}
+\label{fig:jug-deps}
+\end{figure}
+
+This defines the task dependency structure represented in
+Figure~\ref{fig:jug-deps}. As we can see, all of the \code{count} operations
+can be run in parallel, while the \code{mean} operation must wait the result of
+all of the other computation. The dependency structure is always a \textsc{dag}
+(directed acyclic graph). Only by bypassing the normal use mechanisms is it
+even possible to construct a cycle.
+
+The code above has the construct \code{Task(f, args)} repeated several times.
+Using the decorator \code{TaskGenerator} decorator this can be simplified to a
+more natural syntax.
+
+\begin{python}
+from jug import TaskGenerator
+
+@TaskGenerator
+def count(imname):
+    ...
+    return value
+
+@TaskGenerator
+def mean(args):
+    return sum(args)/float(len(args))
+
+images = glob('*.png')
+counts = map(count, images)
+final = mean(counts)
+\end{python}
+
+As the reader can appreciate, this is identical to a traditional Python script,
+except for the \code{@TaskGenerator} decorators.
+
+By default, jug looks for a file called \code{jugfile.py}, but any filename can
+be used. Generically, we refer to the script being run as the jugfile.
+
+\subsection{Jug subcommands}
+
+Jug is structured as a series of subcommands, the most important of which are
+\emph{execute}, \emph{status}, and \emph{shell}.
+
+Execution is the command used for running the tasks. In broad terms it performs
+the loop below.
+
+\begin{python}
+tasks = alltasks in topological sort order
+while tasks:
+    next = task.pop()
+    if not next.has_run() and not next.is_running():
+        while not next.can_run():
+            wait a while
+        with locked(next):
+            next.run()
+\end{python}
+
+If run on a single processor, this will just run all of the tasks in order. It
+is most interesting when it is run on multiple processors. There it executes
+all of the tasks that can be executed in parallel.
+
+Of course, the actual code is more complex than what is shown above,
+particularly to make sure that the locking is performed correctly and that the
+waiting step eventually times out (in order to handle the situation where
+another process is hung).
+
+\begin{figure}
+\begin{verbatim}
+$jug status images.py
+Task name         Waiting      Ready   Finished    Running
+----------------------------------------------------------
+images.mean             1          0          0          0
+images.count            0         20          0          0
+..........................................................
+Total:                  1         20          0          0
+\end{verbatim}
+\caption{Output of \code{jug status}. The \texttt{\$} sign shown is the command
+line prompt, and the status subcommand was run. At this point, nothing has been
+run. The output has been edited for space reasons (spacing columns were
+removed).}
+\label{fig:jug-status-output}
+\end{figure}
+
+The status command prints out a summary of the status of all of the tasks.
+Figure~\ref{fig:jug-status-output} shows the output of this command using the
+example jugfile above. We assume that the jugfile was called \code{images.py}
+on disk and that there were 20~images in the directory. We can see that there
+are 20~tasks ready to run, while the \code{mean} task is still waiting for the
+results of the other tasks.
+
+\subsection{Backends}
+
+A basic feature of jug is its ability to save and load results. Each task
+\code{Task(f, args)} is represented by a hash of \code{f} and \code{args} in a
+way that uniquely identifies it.
+
+A jug backend must then support four basic operations:
+
+\begin{description}
+\item[save] Saving a Python object by its hash name.
+\item[load] Loading a Python object by its hash name.
+\item[lock] Creating a lock by hash name. Naturally, this lock must be created
+atomically.
+\item[release] Releasing the lock.
+\end{description}
+
+A few other operations, such as deletion and listing of names are also
+supported.
+
+The filesystem can support all of the above operations if coded correctly to
+avoid race conditions. This is the default backend, identified simply by a
+directory name. Inside this directory, files named by a hexadecimal
+representation of their hashes. Objects are saved using Python's \code{pickle}
+module with zlib compression. As a special case, \code{numpy} arrays are saved
+to disk directly. For functionality, there was no need for this inelegant
+special case, but \code{numpy} arrays are a very common data type in scientific
+programming and saving them directly allows for very fast saving and loading
+(they are represented on disk as a header followed by the binary information
+they contain).
+
+Another backend currently included with jug is a redis backend. Redis a
+name-key database system.\footnote{See the redis webpage, at
+\href{http://redis.io}{redis.io}, for detailed information about redis.} Redis
+is particularly recommended for the case where there are many small objects
+being saved. In this case, keeping each as a separate file on disk would incur
+a large space penalty, while redis keeps them all in the same file.
+
+Finally, there is an in-memory backend. This was initially developed for
+testing, but can be useful on its own.
+
+Results can be obtained from jug in two ways: (1) One can simply write a task
+that outputs the desired format. (2) They can be inspected interactively using
+the \emph{shell} subcommand.
+
+The shell subcommand, invokes an IPython instance with all the objects in the
+jugfile loaded. IPython is an enhanced interactive shell for Python
+\citep{Perez2007}. A few functions are added to the namespace, in particular,
+\code{value} can load the results of a task object if necessary.
+
+\begin{figure}
+\begin{center}
+\begin{verbatim}
+$jug shell images.py
+
+=========
+Jug Shell
+=========
+
+Available jug functions:
+    - value() : loads a specific object
+    - load_all() : loads all objects
+
+Enjoy...
+
+In [1]: value(final)
+Out[1]: 7.5
+
+In [2]: value(counts)
+Out[2]: [7, 7, 7, 7, 7, 7, 7, 7, 7, 7, 8, 8, 8, 8, 8, 8, 8, 8, 8, 8]
+\end{verbatim}
+\end{center}
+\caption{Interaction with \code{jug shell}.}
+\label{fig:jug-shell-interaction}
+\end{figure}
+
+Figure~\ref{fig:jug-shell-interaction} shows a possible interaction session
+with jug shell. While having to explicitly load all the results is a bit
+bothersome, it has some advantages versus the alternative of pre-loading all.
+It is much faster at startup. Consider too that the user might not load more
+than a few objects. In some cases, loading all of the objects simultaneously
+might even be impossible due to memory constraints. Further, this allows
+exploration of the task structure for debugging.
+
+Jug includes a full test suite. There are no known bugs.
+
+Jug has been available in the Python Package Index\footnote{The Python Package
+Index, PyPI, is accessible at \url{http://pypi.python.org}.} since May~2009
+and has been downloaded over 6000~times (some of these downloads may, of
+course, represent upgrades, and users which downloaded it from some other
+source will not be counted).% FIXME: Check download number. At 5713 8/31/2011
 \section{Methods}

### 2011-11-21 paper.tex
TXT More about development, philosophy, &c

@@
 
+By allowing caching of results, where the caching depends explicitly on the
+input parameters of that computation. Any change in the parameters immediately
+triggers a recomputation of all dependendent results. The basic model is
+similar to the one of the venerable Unix tool Make~\citep{}, which has been
+used before for implementing reproducible research pipelines~\citep{}.
+
+However, unlike make, jug is written in Python and tasks can consist of any
+Python function and its arguments. A jugfile (the file which specifies which
+tasks need to be run---named by analogy to the makefiles of make) is a simple
+Python script with some added notations (we will see just how little changes
+need to be made from a conventional Python script).
+
+\section{Jug}
 \subsection{Task Based Architecture}
@@
 As the reader can appreciate, this is identical to a traditional Python script,
-except for the \code{@TaskGenerator} decorators.
+except for the \code{@TaskGenerator} decorators. With a few limitations (which
+unfortunately can give rise to complex error messages), scripts can be written
+as if they were sequential Python scripts.
 
@@
 
+\section{Methods}
+\section{Example}
+\section{Development of Jug}
+Jug is available under the MIT software license, which grants the user right to
+use, copy, and modify the code almost at will. It is developed using the git
+version control tool, with the project being hosted on the github
+platform.\footnote{The jug project is available at
+\url{https://www.github.com/luispedro/jug}.} An open mailing-list provides
+support and reported bugs are generally fixed within hours.
+
 Jug includes a full test suite. There are no known bugs.
@@
 source will not be counted).% FIXME: Check download number. At 5713 8/31/2011
-\section{Methods}
-\section{Example}
+
 \section{Discussion}
+Unlike most previous tools for reproducible research, jug focuses on the
+development of the computation as much as its communication (in fact, when it
+comes to communication other tools might be better targetted as they combine
+written exposition with computer code). While it still possible to make errors
+(for example, by changing the code but neglecting to invalidate the results),
+jug improves the pipeline development experience and makes the programming
+researcher more productive.
+
+In the last few years, I have not undertaken any computational research project
+without the use of jug to manage my computations. The tool has become an
+indispensible part of my toolkit.
+
 \end{document}

### 2011-11-23 paper.tex
TXT More on the design choices

@@
 
-\section{Jug}
+\section{Design Choices}
 \subsection{Task Based Architecture}
@@
 results of the other tasks.
-
 \subsection{Backends}
@@
 
-The filesystem can support all of the above operations if coded correctly to
-avoid race conditions. This is the default backend, identified simply by a
-directory name. Inside this directory, files named by a hexadecimal
+The filesystem can support all of the above operations if the backend is coded
+correctly to avoid race conditions. This is the default backend, identified
+simply by a directory name. Inside this directory, files named by a hexadecimal
 representation of their hashes. Objects are saved using Python's \code{pickle}
-module with zlib compression. As a special case, \code{numpy} arrays are saved
-to disk directly. For functionality, there was no need for this inelegant
-special case, but \code{numpy} arrays are a very common data type in scientific
-programming and saving them directly allows for very fast saving and loading
-(they are represented on disk as a header followed by the binary information
-they contain).
+module with zlib compression. As a special case, \code{numpy} arrays
+\citep{numpy} are saved to disk directly. For functionality, there was no need
+for this inelegant special case as \code{numpy} can be saved as a pickle dump,
+but \code{numpy} arrays are a very common data type in scientific programming
+and saving them directly allows for very fast saving and loading (they are
+represented on disk as a header followed by the binary information they
+contain).
 
@@
 
-Results can be obtained from jug in two ways: (1) One can simply write a task
-that outputs the desired format. (2) They can be inspected interactively using
-the \emph{shell} subcommand.
+\subsection{Asynchronous Function}
+
+Parallelisation is achieved by running more than one jug process
+simmultaneously. All of the synchronisation is outsourced to the backend. This
+has certain limitations, but it makes jug usable in cluster computing
+environments with a batch architecture. As long as all of the processes can
+access the backend, there is no need for them to communicate directly.
+Therefore, it can even be the case that processors start working on tasks in
+mid-processing.
+
+\subsection{Development Support}
+
+The output of computation can be obtained from jug in two ways: (1) One can
+simply write a task that outputs the desired format. (2) They can be inspected
+interactively using the \emph{shell} subcommand. It is expected that the first
+option will be used for the final version of the computation, whilst the second
+one is most helpful during development.
 
@@
 
+\subsection{Result Invalidation}
+
+It is often the case that one improves a step in a computational pipeline and
+wished to obtain the result of recomputing the pipeline while reusing as much
+of the previously computed intermediate results. With the manual management of
+results I found this to be the most error-prone operation: by not removing
+enough of the intermediate files, it was easy to generate an irreproducible
+state where part of the computations were made using one version of the code
+and part with another.
+
+Therefore, jug adds some support for result invalidation. When a results from a
+task are invalidated, all tasks which, directly or indirectly, depend on them
+are also marked for deletion.
+
+It is still necessary to manually invalidate tasks. An alternative, which I
+considered, was to have the code which implements the task be taken into
+account while computing the hash. While this would add another layer of
+protection, it would still be possible to make mistakes. If the function
+depended on other functions, especially if this was done in a dynamic way, it
+could be hard to discover all dependencies (cf.\ sumatra~\citep{sumatra}).
+Additionally, minute refactoring of the code would lead to over eager
+recomputation. This could make the developer wary of making improvements in
+their code, resulting in overall worse code.
+
+Therefore, I chose to force the user to explicitly invalidate his results,
+while supporting automatic dependency discovery to make it easier for the user.
+
 \section{Methods}

### 2011-11-23 paper.tex
TXT Add example
This is from my own research, so it's a real-life example

@@
 \section{Example}
+I will present an example from my research. Although it introduces some
+unnecessary complexity, it has the advantage of being a real-life example of
+the usage of jug.
+
+The goal of this project was to demonstrate that using local features, in
+particular speeded-up robust features (henceforth,
+\textsc{surf})~\cite{surf-paper} improves over image-global features for
+classification of microscope images. A variation of standard \textsc{surf},
+\textsc{surf}-ref, was introduced and tested.
+
+To evaluate classification, the dataset is broken up into 10~pieces and each of
+the pieces is held-out of the analysis and then used to evaluate analysis (the
+final result is the average of all ten values). This is known as
+cross-validation. In this case, I use another library (which I also wrote)
+which has a builtin method for jug-aware cross-validation.
+
+\begin{python}
+from jug import TaskGenerator, CachedFunction
+
+# pyslic is used for subcellular location analysis
+import pyslic
+
+# milk (machine learning toolkit)
+from milk.ext.jugparallel import nfoldcrossvalidation
+
+datasets = [ 'data0', 'data1', ]
+featuresets = [ 'field-dna+', 'surf', 'surf-ref', ]
+
+def load(dset):
+    '''
+    This function assumes that the images are stored in ../data/dset with
+    filenames matching the pattern
+    label-(protein|dna)-([0-9]).tiff
+    '''
+    from os import listdir
+    images = []
+    base = '../data/'+dset
+    for path in listdir(base):
+        if not 'protein' in path: continue
+        label = path.split('-')[0]
+        # We only store paths and will load data on demand
+        # This save memory.
+        im = pyslic.Image({
+                'protein': base + path,
+                'dna' : base + path.replace('protein','dna')
+                })
+        im.label = label
+        images.append(im)
+    return images
+
+# pyslic.computefeatures takes an image and an identifier of the feature set
+# and returns a set of features (as a numeric array).
+#
+# TaskGenerator can be used directly to adapt existing functions
+computefeatures = TaskGenerator(pyslic.computefeatures)
+
+# or it can be used as a decorator on a newly defined function
+@TaskGenerator
+def output(cmatrix, dset, fset):
+    output = file('results/cmatrix-'+dset+'-'+fset+'.txt', 'w')
+    print >>output, cmatrix
+    output.close()
+
+features = []
+for dset in datasets:
+    images = CachedFunction(load, dset)
+    labels = [im.label for im in images]
+    for fset in featuresets:
+        features = [computefeatures(im, fset) for im in images]
+        cmatrix = nfoldcrossvalidation(features, labels)
+        output(cmatrix, dset, fset)
+\end{python}
+
 \section{Development of Jug}

### 2011-11-24 paper.tex
BLD Add necessary packages

@@
 \documentclass{article}
+\usepackage{hyperref}
+\usepackage{tikz}
+\usepackage{python}
+\usepackage[%
+    natbib=true,%
+    backend=biber,%
+    backref=true,%
+    citestyle=authoryear-comp,%
+    bibstyle=authoryear,%
+    maxbibnames=24,%
+    maxcitenames=2,%
+]{biblatex}
+
+\hypersetup{%
+    bookmarks=true,
+    colorlinks=true,
+    pdfauthor={Luis Pedro Coelho},
+    citecolor=pondgreen,
+    urlcolor=toastedchilipowder,
+    linkcolor=toastedchilipowder,
+}
+\let\code\texttt
 

### 2011-11-28 paper.tex
MIN Minor phrasing fixes

@@
 
-\subsection{Development Support}
+\subsection{Software Development With Jug}
 
@@
 exploration of the task structure for debugging.
+
+This is a feature which can be used to debug one's own code or to explore the
+results of a computation by others.
 
@@
 development of the computation as much as its communication (in fact, when it
-comes to communication other tools might be better targetted as they combine
+comes to communication other tools might be better suited as they combine
 written exposition with computer code). While it still possible to make errors

### 2011-11-29 paper.tex
BIB A few references

@@
 \begin{document}
+\maketitle
+
 \section*{Abstract}
@@
 \url{http://github.com/luispedro/jug}.
+\bigskip
+\bigskip
 
@@
 triggers a recomputation of all dependendent results. The basic model is
-similar to the one of the venerable Unix tool Make~\citep{}, which has been
-used before for implementing reproducible research pipelines~\citep{}.
+similar to the one of the venerable Unix tool Make, which has been used before
+for implementing reproducible research
+pipelines~\citep{Schwab00makingscientific}.
 

### 2011-11-30 paper.tex
MIN Better phrasing

@@
 simmultaneously. All of the synchronisation is outsourced to the backend. This
-has certain limitations, but it makes jug usable in cluster computing
+has certain limitations, but it makes jug usable in certain cluster computing
 environments with a batch architecture. As long as all of the processes can
-access the backend, there is no need for them to communicate directly.
-Therefore, it can even be the case that processors start working on tasks in
-mid-processing.
+access the backend, there is no need for them to communicate directly. It can
+even be the case that processors start working on tasks in mid-processing.
 

### 2011-11-30 paper.tex
TXT+BIB Add new ref. Compare to it

@@
 need to be made from a conventional Python script).
+
+\citet{Fomel2007} proposed a system built on top of Scons which shared
+superficially similar design. Scons, like Make, is a build system. It supports
+spawning parallel jobs using multiple threads. In the original use case, the
+tasks are delegated to other commands and the operating system can take care of
+parallelism. If, however, the tasks are computationally intensive in Python,
+contention for the global interpreter lock will limit the amount of real
+parallelism.\footnote{In the most commonly used Python interpreters, there is a
+lock that prevents more than one thread from simultaneously executing
+interpreted code, although they can execute non-interpreted code, as calling an
+external programme or certain external libraries.}
 
@@
 with jug shell. While having to explicitly load all the results is a bit
-bothersome, it has some advantages versus the alternative of pre-loading all.
-It is much faster at startup. Consider too that the user might not load more
-than a few objects. In some cases, loading all of the objects simultaneously
-might even be impossible due to memory constraints. Further, this allows
-exploration of the task structure for debugging.
+bothersome, it can be better than pre-loading all of them. It is much faster at
+startup. Consider too that the user might not load more than a few objects. In
+some cases, loading all of the objects simultaneously might even be impossible
+due to memory constraints. Further, this allows exploration of the task
+structure for debugging.
 

### 2011-11-30 paper.tex
TXT Update download count

@@
 Index, PyPI, is accessible at \url{http://pypi.python.org}.} since May~2009
-and has been downloaded over 6000~times (some of these downloads may, of
+and has been downloaded over 7000~times (some of these downloads may, of
 course, represent upgrades, and users which downloaded it from some other
-source will not be counted).% FIXME: Check download number. At 5713 8/31/2011
+source will not be counted).% FIXME: Check download number. At 7484 11/29/2011
 

### 2011-12-08 paper.tex
MIN Add link to previous paper's archives
MIN Update download count
A few downloads since last release

--- 2011-11-30 paper.tex
@@
 Index, PyPI, is accessible at \url{http://pypi.python.org}.} since May~2009
-and has been downloaded over 6000~times (some of these downloads may, of
+and has been downloaded over 8000~times (some of these downloads may, of
 course, represent upgrades, and users which downloaded it from some other
-source will not be counted).% FIXME: Check download number. At 5713 8/31/2011
+source will not be counted).% FIXME: Check download number. At 8112 12/8/2011
 
@@
 In the last few years, I have not undertaken any computational research project
-without the use of jug to manage my computations. The tool has become an
-indispensible part of my toolkit.
+without the use of jug to manage my computations. Additionally, I (and my
+co-authors) have made the resulting results directories available\footnote{See
+the links at
+\url{http://murphylab.cbi.cmu.edu/software/software_from_papers.html}}. The
+tool has become an indispensible part of my toolkit.
 

### 2011-12-08 paper.tex
TXT Add fan-in/fan-out. Discuss Hadoop

@@
 
+Figure~\ref{fig:jug-deps-complex} shows the dependency structure of this
+examples. The main feature is the fan-out/fan-in which is associated with
+map-reduce frameworks.
+
+\begin{figure}
+\begin{center}
+\begin{tikzpicture}
+\node (load) at (0,0) [draw] {load};
+\node (collect) at (4,0) [draw] {collect};
+\node (sum) at (8,0) [draw] {sum};
+\node (output) at (10,0) [draw] {output};
+
+\path[->]
+    (sum) edge (output);
+
+\foreach \y in {-2,-1,0,1,2} {%
+    \node (features\y) at (2,\y) [draw] {features};
+    \path[->]
+        (load) edge (features\y)
+        (features\y) edge (collect);
+}
+
+\foreach \y in {-2,-1,0,1,2} {%
+    \node (fold\y) at (6,\y) [draw] {fold};
+    \path[->]
+        (collect) edge (fold\y)
+        (fold\y) edge (sum);
+}
+\end{tikzpicture}
+
+\end{center}
+\caption{Dependency Structure for second example in the Text. The
+fan-out/fan-in structure is typical (it is an instance of the map-reduce
+framework).}
+\label{fig:jug-deps-complex}
+\end{figure}
+
 \section{Development of Jug}
@@
 
+Jug was not designed to compete with more complex map/reduce frameworks such as
+Hadoop, which scale better to very large projects. However, those frameworks
+have much higher costs in terms of development time (code must be written
+especially for the framework) and overhead (starting up and shutting down jobs
+can take much longer and using non-Java code has additional computational
+costs). For scientists using small to medium-sized cluster (hundreds, not
+thousands, of nodes), with a shared file-system (as opposed to the distributed
+file-systems necessary for truly large amounts of data) jug might provide a
+better trade-off  than those higher-powered alternatives.
+
 In the last few years, I have not undertaken any computational research project

### 2012-01-23 paper.tex
TEX Add bibliography
STYL Make it narrower

@@
 \let\code\texttt
+\addbibresource{references.bib}
 
@@
 In [2]: value(counts)
-Out[2]: [7, 7, 7, 7, 7, 7, 7, 7, 7, 7, 8, 8, 8, 8, 8, 8, 8, 8, 8, 8]
+Out[2]: [7, 7, 7, 7, 7, ... 8, 8, 8, 8, 8, 8, 8, 8, 8]
 \end{verbatim}
 \end{center}
-\caption{Interaction with \code{jug shell}.}
+\caption{Interaction with \code{jug shell}. The \code{value} function loads and
+returns the result from any task. The final line has been edited (marked with
+\ldots) for presentation purposes.}
 \label{fig:jug-shell-interaction}
@@
     '''
-    This function assumes that the images are stored in ../data/dset with
-    filenames matching the pattern
+    This function assumes that the images are stored in
+    ../data/dset with filenames matching the pattern
     label-(protein|dna)-([0-9]).tiff
@@
 
-# pyslic.computefeatures takes an image and an identifier of the feature set
-# and returns a set of features (as a numeric array).
+# pyslic.computefeatures takes an image and an identifier of the
+# feature set and returns a set of features (as a numeric array).
 #
@@
 
+\printbibliography
 \end{document}

### 2012-01-23 paper.tex
ENH Comments about sumatra.
I did not find a proper paper for it, so I added a one-sentence
explanation and a link to its webpage.

@@
 depended on other functions, especially if this was done in a dynamic way, it
-could be hard to discover all dependencies (cf.\ sumatra~\citep{sumatra}).
-Additionally, minute refactoring of the code would lead to over eager
-recomputation. This could make the developer wary of making improvements in
-their code, resulting in overall worse code.
+could be hard to discover all dependencies (cf.\ sumatra~\footnote{Sumatra is a
+tool to collect all dependencies that are necessary to run a particular
+computation. Its webpage is located at
+\url{http://neuralensemble.org/trac/sumatra}.}). Additionally, minute
+refactoring of the code would lead to over eager recomputation. This could make
+the developer wary of making improvements in their code, resulting in overall
+worse code.
 
@@
 while supporting automatic dependency discovery to make it easier for the user.
+The recommendation is still that the user run the full pipeline from start to
+finish once they are satisfied with the state of the code and certainly before
+publication.
 

### 2012-01-24 paper.tex
TXT Comments on splitting computation/figures
Often, the problem of "this computation takes a few computer years" is
not properly acknowledged.

@@
 Unlike most previous tools for reproducible research, jug focuses on the
-development of the computation as much as its communication (in fact, when it
+development of the computation as much as its communication. In fact, when it
 comes to communication other tools might be better suited as they combine
-written exposition with computer code). While it still possible to make errors
-(for example, by changing the code but neglecting to invalidate the results),
-jug improves the pipeline development experience and makes the programming
-researcher more productive.
+written exposition with computer code. They can be used in combination with jug
+as they do not often provide caching and distribution, needed for large
+projects. While the simplest mode of jug use is to have a single
+\textit{jugfile}, a jugfile can import the results of another. For example, a
+first jugfile might perform all heavy-duty computation and be run on a
+computing cluster. A second script could be embedded inside an executable paper
+to generate tables and plots \citep{Delescluse2011}. This faster script could
+be run as part of building the paper (while the longer script would take too
+long).
 
-Jug was not designed to compete with more complex map/reduce frameworks such as
+While it still possible to make errors (for example, by changing the code but
+neglecting to invalidate the results), jug improves the pipeline development
+experience and makes the programming researcher more productive and less
+error-prone.
+
+Jug was not designed to compete with complete map/reduce frameworks such as
 Hadoop, which scale better to very large projects. However, those frameworks

### 2012-01-25 paper.tex
TXT Better introduction

@@
 \section{Introduction}
+The value of reproducible research in computational fields has been recognized
+in several areas \citep{Vandewalle2009,Nordlie2009}. Although some of the
+reports are still more forward looking rather than descriptive of current
+practice, several implementations of reproducible papers (or executable papers)
+have been proposed towards the goal of reproducing published analyses.
+
+These solutions do not necessarily scale to large problems, those that take
+several days, months, or years of CPU time. For very large problems,
+specialized solutions are needed, but there is a range of medium-sized problems
+that can be successfully tackled on a computer cluster with a few hundred nodes
+or less (so that a computation that would take a CPU-year can produce results
+in less than a day).
+
+The typical ad hoc solution to this problem is to save intermediate files.
+Often the design of the computation itself takes several iterations as
+intermediate steps are improved. Thus, some intermediate results need to be
+recomputed. This involves a great amount of human management of the state of
+the computation, breaking it up into pieces, and issuing jobs on the cluster.
+
 Jug is a task-based framework for Python, which supports saving and sharing of

### 2012-02-23 paper.tex
TXT Cite similar tools

@@
 interpreted code, although they can execute non-interpreted code, as calling an
-external programme or certain external libraries.}
+external programme or certain external libraries.} Ruffus is another
+Python-based solution which supports parallel execution of pipelines
+\citep{Goodstadt16092010}. There is a strong dependency on the filesystem,
+which makes its use very different from jug usage, which strives to make its
+usage similar to straight Python usage. Using the Ruby programming language,
+\citet{mishima2011} proposed Pwrake.
 

### 2012-02-23 paper.tex
TXT Better text

@@
 usage similar to straight Python usage. Using the Ruby programming language,
-\citet{mishima2011} proposed Pwrake.
+\citet{mishima2011} proposed Pwrake, which supports parallel execution of
+bioinformatics workflows written (or accessible) in that language.
 
@@
 Jug is designed around tasks. A task is defined as a Python function and a set
-of arguments, which may be Python values or the output of other tasks. A task
-should be a pure function, but there is no way of mandating this in the
-language.
+of arguments, which may be Python values or the output of other tasks.
 
@@
 
-This defines the task dependency structure represented in
+This code defines the task dependency structure represented in
 Figure~\ref{fig:jug-deps}. As we can see, all of the \code{count} operations
@@
 all of the other computation. The dependency structure is always a \textsc{dag}
-(directed acyclic graph). Only by bypassing the normal use mechanisms is it
-even possible to construct a cycle.
+(directed acyclic graph).
 
@@
 
+This makes jug usable in batch-based computer cluster environments, which are
+quite common in research institutions.
+
 \subsection{Software Development With Jug}
@@
 The output of computation can be obtained from jug in two ways: (1) One can
-simply write a task that outputs the desired format. (2) They can be inspected
-interactively using the \emph{shell} subcommand. It is expected that the first
-option will be used for the final version of the computation, whilst the second
-one is most helpful during development.
+simply write a task that writes to a file in the desired format. (2) They can
+be inspected interactively using the \emph{shell} subcommand. It is expected
+that the first option will be used for the final version of the computation,
+whilst the second one is most helpful during development.
 
@@
 finish once they are satisfied with the state of the code and certainly before
-publication.
+publication, but the pipeline development stage can be more agile.
 
@@
 Index, PyPI, is accessible at \url{http://pypi.python.org}.} since May~2009
-and has been downloaded over 8000~times (some of these downloads may, of
-course, represent upgrades, and users which downloaded it from some other
-source will not be counted).% FIXME: Check download number. At 8112 12/8/2011
+through February~2012 and has been downloaded over 10,000~times (some of these
+downloads may, of course, represent upgrades, and users which downloaded it
+from some other source will not be counted).
 

### 2012-04-25 paper.tex
TXT Contextualize jug wrt incpy

@@
 need to be made from a conventional Python script).
-
-\citet{Fomel2007} proposed a system built on top of Scons which shared
-superficially similar design. Scons, like Make, is a build system. It supports
-spawning parallel jobs using multiple threads. In the original use case, the
-tasks are delegated to other commands and the operating system can take care of
-parallelism. If, however, the tasks are computationally intensive in Python,
-contention for the global interpreter lock will limit the amount of real
-parallelism.\footnote{In the most commonly used Python interpreters, there is a
-lock that prevents more than one thread from simultaneously executing
-interpreted code, although they can execute non-interpreted code, as calling an
-external programme or certain external libraries.} Ruffus is another
-Python-based solution which supports parallel execution of pipelines
-\citep{Goodstadt16092010}. There is a strong dependency on the filesystem,
-which makes its use very different from jug usage, which strives to make its
-usage similar to straight Python usage. Using the Ruby programming language,
-\citet{mishima2011} proposed Pwrake, which supports parallel execution of
-bioinformatics workflows written (or accessible) in that language.
-
 \section{Design Choices}
@@
 \section{Discussion}
+
+\subsection{Similar Tools}
+
+\citet{Fomel2007} proposed a system built on top of Scons which shared
+superficially similar design. Scons, like Make, is a build system. It supports
+spawning parallel jobs using multiple threads. In the original use case, the
+tasks are delegated to other commands and the operating system can take care of
+parallelism. If, however, the tasks are computationally intensive in Python,
+contention for the global interpreter lock will limit the amount of real
+parallelism.\footnote{In the most commonly used Python interpreters, there is a
+lock that prevents more than one thread from simultaneously executing
+interpreted code, although they can execute non-interpreted code, as calling an
+external programme or certain external libraries.} Ruffus is another
+Python-based solution which supports parallel execution of pipelines
+\citep{Goodstadt16092010}. There is a strong dependency on the filesystem,
+which makes its use very different from jug usage, which strives to make its
+usage similar to straight Python usage. Using the Ruby programming language,
+\citet{mishima2011} proposed Pwrake, which supports parallel execution of
+bioinformatics workflows written (or accessible) in that language.
+
+Of interest, is also the \textsc{IncPy} implementation of a Python interpreter
+with automatic persistent memoization~\cite{Guo_IncPy,Guo_IncPy_pre}. The
+arguments made for that system apply to Jug as well, with a few differences.
+Their system implements automatic memoization, while, using Jug, the user needs
+to automatically annotate functions to memoize using \texttt{TaskGenerator}.
+This extra control can be necessary to avoid a proliferation of intermediate
+results when these are very large (often a very large intermediate output is
+not necessary, and only a summary must be kept). Additionally, Jug supports
+running tasks in parallel, a functionality that is absent from \textsc{IncPy}.
+
+\subsection{Conclusions}
+
 Unlike most previous tools for reproducible research, jug focuses on the

### 2012-06-08 paper.tex
ENH Fix typos & improve text

@@
 
-In addition to the value to the community of being able to reproduce an
-analysis, reproducible research practices allow for better control over the
-project. For example, if necessary parameters to run a pipeline are kept
-separately from the code that implements it (perhaps even in the researcher's
-mind), this leads to error-prone analysis and opens up the possibility that
-when the results are to be writtne up for publication, the researcher will no
-longer be able to even completely describe the process that led to them.
+In addition to being valuable to the wider community, which may wish to
+reproduce or extend a published analysis, reproducible research practices allow
+for better control over the project by the authors themselves. For example, if
+necessary parameters to run a pipeline are kept separately from the code that
+implements it (perhaps even in mind of the researcher), this leads to
+error-prone analysis and opens up the possibility that when the results are to
+be written up for publication, the researcher will no longer be able to even
+completely describe the process that led to them.
 
@@
 or distributed across a cluster) is necessary to obtain results in a useful
-timeframe. Furthermore, it is often the case that, as the project evolves, it
+time frame. Furthermore, it is often the case that, as the project evolves, it
 becomes necessary to save intermediate results while down-stream analyses are
-designed (or redesigned) and implemented. Therefore, having a single point of
-entry (often in the form of a single script or a single main programme) for the
-computation becomes increasingly difficult.
+designed (or redesigned) and implemented. Under many frameworks, this causes
+having a single point of entry (often in the form of a single script or a
+single main programme) for the computation becomes increasingly difficult.
 
-Jug is a software framework which solves all of these problems in a
-simple way. Jug supports caching of intermediate results, distribution of
-computation as tasks across a network.
+Jug is a software framework which solves all of these problems in a simple way.
+Jug supports caching of intermediate results, distribution of computation as
+tasks across a network.
 
 Jug is written in pure Python, is completely cross-platform, and available as
-free software under the MIT license. Jug is available from
+free software under the liberal MIT license. Jug is available from
 \url{http://github.com/luispedro/jug}.
@@
 several days, months, or years of CPU time. For very large problems,
-specialized solutions are needed, but there is a range of medium-sized problems
-that can be successfully tackled on a computer cluster with a few hundred nodes
-or less (so that a computation that would take a CPU-year can produce results
-in less than a day).
+specialized solutions are needed~\citep{mapReduce}. There is a range of
+medium-sized problems that can be successfully tackled on a computer cluster
+with a few hundred nodes or less (so that a computation that would take a
+CPU-year can produce results in less than a day).
 
@@
 input parameters of that computation. Any change in the parameters immediately
-triggers a recomputation of all dependendent results. The basic model is
-similar to the one of the venerable Unix tool Make, which has been used before
-for implementing reproducible research
-pipelines~\citep{Schwab00makingscientific}.
+triggers a recomputation of all dependent results. The basic model is similar
+to the one of the venerable Unix tool Make, which has been used before for
+implementing reproducible research pipelines~\citep{Schwab00makingscientific}.
 
 However, unlike make, jug is written in Python and tasks can consist of any
-Python function and its arguments. A jugfile (the file which specifies which
-tasks need to be run---named by analogy to the makefiles of make) is a simple
-Python script with some added notations (we will see just how little changes
-need to be made from a conventional Python script).
+Python function and its arguments. This can include running external commands.
+A jugfile (the file which specifies which tasks need to be run---named by
+analogy to the makefiles of make) is a simple Python script with some added
+notations (we will see just how little changes need to be made from a
+conventional Python script).
+
 \section{Design Choices}
@@
 
-Execution is the command used for running the tasks. In broad terms it performs
-the loop below.
+Execution is the command used for running the tasks. It performs a more complex
+version of this loop:
 
@@
 results of the other tasks.
+
 \subsection{Backends}
@@
 Parallelisation is achieved by running more than one jug process
-simmultaneously. All of the synchronisation is outsourced to the backend. This
-has certain limitations, but it makes jug usable in certain cluster computing
-environments with a batch architecture. As long as all of the processes can
-access the backend, there is no need for them to communicate directly. It can
-even be the case that processors start working on tasks in mid-processing.
-
-This makes jug usable in batch-based computer cluster environments, which are
-quite common in research institutions.
+simultaneously. All of the synchronisation is outsourced to the backend. As
+long as all of the processes can access the backend, there is no need for them
+to communicate directly. It can even be the case that processors start working
+on tasks in mid-processing. This makes jug usable in batch-based computer
+cluster environments, which are quite common in research institutions.
 
@@
 wished to obtain the result of recomputing the pipeline while reusing as much
-of the previously computed intermediate results. With the manual management of
-results I found this to be the most error-prone operation: by not removing
-enough of the intermediate files, it was easy to generate an irreproducible
-state where part of the computations were made using one version of the code
-and part with another.
+of the previously computed intermediate results. Before using Jug, I found this
+to be a very error-prone operation: by not removing enough of the intermediate
+files, it was easy to generate an irreproducible state where part of the
+computations were made using one version of the code and part with another.
 
@@
 computation. Its webpage is located at
-\url{http://neuralensemble.org/trac/sumatra}.}). Additionally, minute
+\url{http://neuralensemble.org/trac/sumatra}.}). Additionally, even minute
 refactoring of the code would lead to over eager recomputation. This could make
@@
 cross-validation. In this case, I use another library (which I also wrote)
-which has a builtin method for jug-aware cross-validation.
+which has a built-in method for jug-aware cross-validation.
 
@@
 Index, PyPI, is accessible at \url{http://pypi.python.org}.} since May~2009
-through February~2012 and has been downloaded over 10,000~times (some of these
+through June~2012 and has been downloaded over 13,000~times (some of these
 downloads may, of course, represent upgrades, and users which downloaded it

### 2013-05-16 paper.tex
TXT Update download count.
I wonder if the previous very low value was a typo.

@@
 Index, PyPI, is accessible at \url{http://pypi.python.org}.} since May~2009
-through June~2012 and has been downloaded over 13,000~times (some of these
+through May~2013 and has been downloaded over 43,000~times (some of these
 downloads may, of course, represent upgrades, and users which downloaded it

### 2016-10-14 paper.tex
MIN Improve text

@@
 
-\title{Jug: Software for Reproducible Computation}
+\title{Jug: Software for Reproducible Computation In Python}
 \author{Luis Pedro Coelho}
@@
 necessary parameters to run a pipeline are kept separately from the code that
-implements it (perhaps even in mind of the researcher), this leads to
+implements it (in some instance even in mind of the researcher), this leads to
 error-prone analysis and opens up the possibility that when the results are to
@@
 However, unlike make, jug is written in Python and tasks can consist of any
-Python function and its arguments. This can include running external commands.
+Python function and its arguments. This can include running external commands
+and calling routines written in other languages as Python has many tools to
+interface with the rest of the system.
+
 A jugfile (the file which specifies which tasks need to be run---named by
 analogy to the makefiles of make) is a simple Python script with some added
-notations (we will see just how little changes need to be made from a
-conventional Python script).
+notations (we will see the only a few, small, changes are needed to transform a
+conventional Python script into a Jugfile).
 
@@
 By default, jug looks for a file called \code{jugfile.py}, but any filename can
-be used. Generically, we refer to the script being run as the jugfile.
+be used. Generically, we refer to the script being run as the Jugfile.
 
@@
 If run on a single processor, this will just run all of the tasks in order. It
-is most interesting when it is run on multiple processors. There it executes
-all of the tasks that can be executed in parallel.
+is most interesting when it is run on multiple processors. Because of the lock
+synchronisation, tasks can be run in parallel to take advantage of the multiple
+cores.
 
-Of course, the actual code is more complex than what is shown above,
-particularly to make sure that the locking is performed correctly and that the
-waiting step eventually times out (in order to handle the situation where
-another process is hung).
+The actual code is more complex than what is shown above, particularly to make
+sure that the locking is performed correctly and that the waiting step
+eventually times out (in order to handle the situation where another process is
+hung).
 
@@
 module with zlib compression. As a special case, \code{numpy} arrays
-\citep{numpy} are saved to disk directly. For functionality, there was no need
-for this inelegant special case as \code{numpy} can be saved as a pickle dump,
-but \code{numpy} arrays are a very common data type in scientific programming
-and saving them directly allows for very fast saving and loading (they are
+\citep{numpy} are saved to disk directly. This special case was introduced as
+\code{numpy} arrays are a very common data type in scientific programming and
+saving them directly allows for very fast saving and loading (they are
 represented on disk as a header followed by the binary information they
@@
 
-The output of computation can be obtained from jug in two ways: (1) One can
-simply write a task that writes to a file in the desired format. (2) They can
-be inspected interactively using the \emph{shell} subcommand. It is expected
-that the first option will be used for the final version of the computation,
-whilst the second one is most helpful during development.
+The output of computation can be obtained from jug in three ways: (1) One can
+simply write a task that writes the output to a file in the desired format. (2)
+Outputs can be inspected interactively using the \emph{shell} subcommand. It is
+expected that the first option will be used for the final version of the
+computation, whilst the second one is most helpful during development. Finally,
+(3) jug can be loaded as a library from a Python script and jug computation
+outputs can serve as inputs for further computation.
 
@@
 
-It is often the case that one improves a step in a computational pipeline and
-wished to obtain the result of recomputing the pipeline while reusing as much
-of the previously computed intermediate results. Before using Jug, I found this
-to be a very error-prone operation: by not removing enough of the intermediate
-files, it was easy to generate an irreproducible state where part of the
-computations were made using one version of the code and part with another.
+If a developer improves an intermediate step in a pipeline (e.g., fixes a bug)
+and wishes to obtain new results of the computation, then all outputs from that
+step and downstream must be recomputed, but results from upstream and unrelated
+processes can be reused. Formally, in the task DAG, affected tasks and their
+descendants must be recomputed. Without tool support, this can be a very
+error-prone operation: by not removing enough of the intermediate files, it was
+easy to generate an irreproducible state where part of the computations were
+made using one version of the code and part with another.
 
-Therefore, jug adds some support for result invalidation. When a results from a
-task are invalidated, all tasks which, directly or indirectly, depend on them
-are also marked for deletion.
+Therefore, jug adds support for result invalidation. When a results from a task
+are invalidated, all tasks which, directly or indirectly, depend on them are
+also marked for deletion.
 
-It is still necessary to manually invalidate tasks. An alternative, which I
-considered, was to have the code which implements the task be taken into
-account while computing the hash. While this would add another layer of
-protection, it would still be possible to make mistakes. If the function
-depended on other functions, especially if this was done in a dynamic way, it
-could be hard to discover all dependencies (cf.\ sumatra~\footnote{Sumatra is a
-tool to collect all dependencies that are necessary to run a particular
-computation. Its webpage is located at
-\url{http://neuralensemble.org/trac/sumatra}.}). Additionally, even minute
-refactoring of the code would lead to over eager recomputation. This could make
-the developer wary of making improvements in their code, resulting in overall
-worse code.
+It is still necessary to manually invalidate tasks. An alternative, was to have
+the code which implements the task be taken into account while computing the
+hash. While this would add another layer of protection, it would still be
+possible to make mistakes. If the function depended on other functions,
+especially if this was done in a dynamic way, it could be hard to discover all
+dependencies (cf.\ sumatra~\footnote{Sumatra is a tool to collect all
+dependencies that are necessary to run a particular computation. Its webpage is
+located at \url{http://neuralensemble.org/trac/sumatra}.}). Additionally, even
+minute refactoring of the code would lead to over eager recomputation. This
+could make the developer wary of making improvements in their code, resulting
+in overall worse code.
 
-Therefore, I chose to force the user to explicitly invalidate his results,
-while supporting automatic dependency discovery to make it easier for the user.
-The recommendation is still that the user run the full pipeline from start to
-finish once they are satisfied with the state of the code and certainly before
-publication, but the pipeline development stage can be more agile.
+Therefore, as a design choice, jug asks the user to explicitly invalidate their
+results, while supporting automatic dependency discovery. The recommendation is
+still that the user run the full pipeline from start to finish once they are
+satisfied with the state of the code and certainly before publication, but the
+pipeline development stage can be more agile.
 
@@
 \section{Example}
+
 I will present an example from my research. Although it introduces some
@@
 
-Jug has been available in the Python Package Index\footnote{The Python Package
-Index, PyPI, is accessible at \url{http://pypi.python.org}.} since May~2009
-through May~2013 and has been downloaded over 43,000~times (some of these
-downloads may, of course, represent upgrades, and users which downloaded it
-from some other source will not be counted).
-
 \section{Discussion}
@@
 
-While it still possible to make errors (for example, by changing the code but
-neglecting to invalidate the results), jug improves the pipeline development
-experience and makes the programming researcher more productive and less
-error-prone.
+Jug improves the pipeline development experience and makes the programming
+researcher more productive and less error-prone. As it is based on Python, a
+widely used scientific programming language, it can be used in a known
+environment without a long learning process.
 
-Jug was not designed to compete with complete map/reduce frameworks such as
-Hadoop, which scale better to very large projects. However, those frameworks
+Jug was not designed to compete with large scale frameworks such as Hadoop or
+Spark, which scale better to very large projects. However, those frameworks
 have much higher costs in terms of development time (code must be written
-especially for the framework) and overhead (starting up and shutting down jobs
-can take much longer and using non-Java code has additional computational
-costs). For scientists using small to medium-sized cluster (hundreds, not
-thousands, of nodes), with a shared file-system (as opposed to the distributed
-file-systems necessary for truly large amounts of data) jug might provide a
-better trade-off than those higher-powered alternatives.
-
-In the last few years, I have not undertaken any computational research project
-without the use of jug to manage my computations. Additionally, I (and my
-co-authors) have made the resulting results directories available\footnote{See
-the links at
-\url{http://murphylab.cbi.cmu.edu/software/software_from_papers.html}}. The
-tool has become an indispensible part of my toolkit.
+especially for the framework) and overhead. For scientists using small to
+medium-sized cluster (hundreds, not thousands, of nodes), with a shared
+file-system (as opposed to the distributed file-systems necessary for truly
+large amounts of data) jug provides a better trade-off than those
+higher-powered alternatives in a familiar programming language.
 

### 2016-10-29 paper.tex
ENH Minor text improvements

@@
 As computational pipelines become a bigger part of science, it is important to
-ensure that the results are reproducible. All developed software should be able
-to be run automatically without any user intervention.
+ensure that the results are reproducible, a concern which has come to the fore
+in recent years. All developed software should be able to be run automatically
+without any user intervention.
 
@@
 necessary parameters to run a pipeline are kept separately from the code that
-implements it (in some instance even in mind of the researcher), this leads to
-error-prone analysis and opens up the possibility that when the results are to
-be written up for publication, the researcher will no longer be able to even
-completely describe the process that led to them.
+implements it, this leads to error-prone analysis and opens up the possibility
+that when the results are to be written up for publication, the researcher will
+no longer be able to even completely describe the process that led to them.
 
-For large projects, the use of multiple processors (either in the same machine
-or distributed across a cluster) is necessary to obtain results in a useful
-time frame. Furthermore, it is often the case that, as the project evolves, it
-becomes necessary to save intermediate results while down-stream analyses are
-designed (or redesigned) and implemented. Under many frameworks, this causes
-having a single point of entry (often in the form of a single script or a
-single main programme) for the computation becomes increasingly difficult.
+For large projects, the use of multiple computational ccores (either in the
+same machine or distributed across a cluster) is necessary to obtain results in
+a useful time frame. Furthermore, it is often the case that, as the project
+evolves, it becomes necessary to save intermediate results while down-stream
+analyses are designed (or redesigned) and implemented. Under many frameworks,
+this causes having a single point of entry (often in the form of a single
+script or a single main programme) for the computation becomes increasingly
+difficult.
 
@@
 
-The typical ad hoc solution to this problem is to save intermediate files.
-Often the design of the computation itself takes several iterations as
+The typical ad hoc solution to this problem is to save intermediate files on
+disk. Often the design of the computation itself takes several iterations as
 intermediate steps are improved. Thus, some intermediate results need to be
@@
 of arguments, which may be Python values or the output of other tasks.
+
+For example:
 
@@
 Execution is the command used for running the tasks. It performs a more complex
-version of this loop:
+version of the following pseudo-code:
 
@@
 
-The status command prints out a summary of the status of all of the tasks.
-Figure~\ref{fig:jug-status-output} shows the output of this command using the
-example jugfile above. We assume that the jugfile was called \code{images.py}
-on disk and that there were 20~images in the directory. We can see that there
-are 20~tasks ready to run, while the \code{mean} task is still waiting for the
-results of the other tasks.
+The \texttt{status} subcommand prints out a summary of the status of all of the
+tasks. Figure~\ref{fig:jug-status-output} shows the output of this command
+using the example jugfile above. We assume that the jugfile was called
+\code{images.py} on disk and that there were 20~images in the directory. We can
+see that there are 20~tasks ready to run, while the \code{mean} task is still
+waiting for the results of the other tasks.
 
@@
 
-It is still necessary to manually invalidate tasks. An alternative, was to have
-the code which implements the task be taken into account while computing the
-hash. While this would add another layer of protection, it would still be
-possible to make mistakes. If the function depended on other functions,
-especially if this was done in a dynamic way, it could be hard to discover all
-dependencies (cf.\ sumatra~\footnote{Sumatra is a tool to collect all
-dependencies that are necessary to run a particular computation. Its webpage is
-located at \url{http://neuralensemble.org/trac/sumatra}.}). Additionally, even
-minute refactoring of the code would lead to over eager recomputation. This
-could make the developer wary of making improvements in their code, resulting
-in overall worse code.
+It is still necessary to manually invalidate tasks. An alternative, would be to
+take code which implements the task into account while computing the hash. This
+would mark any results computed with this code as outdated if the code changed.
+While this would add another layer of protection, it would still be possible to
+make mistakes. If the function depended on other functions, especially if this
+was done in a dynamic way, it could be hard to discover all dependencies (cf.\
+sumatra~\footnote{Sumatra is a tool to collect all dependencies that are
+necessary to run a particular computation. Its webpage is located at
+\url{http://neuralensemble.org/trac/sumatra}.}). Additionally, even minute
+refactoring of the code would lead to over eager recomputation. This could make
+the developer wary of making improvements in their code, resulting in overall
+worse code.
 

### 2016-10-29 build.sh
MIN Use /usr/bin/env to get zsh path

@@ -1,4 +1,4 @@
-#!/usr/bin/zsh
+#!/usr/bin/env zsh
 
 function build() {
   input=$1

### 2016-11-12 paper.tex
ENH Update paper
Back several assertions with appropriate cites
Improve 'related tools' section

@@
 \addbibresource{references.bib}
+\addbibresource{readcube_export.bib}
 
@@
 reproduce or extend a published analysis, reproducible research practices allow
-for better control over the project by the authors themselves. For example, if
-necessary parameters to run a pipeline are kept separately from the code that
-implements it, this leads to error-prone analysis and opens up the possibility
-that when the results are to be written up for publication, the researcher will
-no longer be able to even completely describe the process that led to them.
+for better control over the project by the original authors themselves. For
+example, if necessary parameters to run a pipeline are kept separately from the
+code that implements it, this leads to error-prone analysis and opens up the
+possibility that, when the results are to be written up for publication, the
+researcher will no longer be able to even completely describe the process that
+led to them.
 
-For large projects, the use of multiple computational ccores (either in the
-same machine or distributed across a cluster) is necessary to obtain results in
-a useful time frame. Furthermore, it is often the case that, as the project
-evolves, it becomes necessary to save intermediate results while down-stream
-analyses are designed (or redesigned) and implemented. Under many frameworks,
-this causes having a single point of entry (often in the form of a single
-script or a single main programme) for the computation becomes increasingly
-difficult.
+For large projects, the use of multiple computational cores (either in a
+multi-core machine or distributed across a compute cluster) is necessary to
+obtain results in a useful time frame. Furthermore, it is often the case that,
+as the project evolves, it becomes necessary to save intermediate results while
+down-stream analyses are designed (or re-designed) and implemented. Under many
+frameworks, this causes having a single point of entry (often in the form of a
+single script or a single main programme) for the computation becomes
+increasingly difficult.
 
@@
 The value of reproducible research in computational fields has been recognized
-in several areas \citep{Vandewalle2009,Nordlie2009}. Although some of the
-reports are still more forward looking rather than descriptive of current
-practice, several implementations of reproducible papers (or executable papers)
-have been proposed towards the goal of reproducing published analyses.
+in several areas
+\citep{donoho2009reproducible,v2009reproducible,nordlie2009towards,fomel2015reproducible,leveque2012reproducible,dudley2010reproducible}.
+Although some of the reports are still more forward looking rather than
+descriptive of current practice, several implementations of reproducible papers
+(or executable papers) have been proposed towards the goal of reproducing
+published analyses.
 
@@
 several days, months, or years of CPU time. For very large problems,
-specialized solutions are needed~\citep{mapReduce}. There is a range of
-medium-sized problems that can be successfully tackled on a computer cluster
-with a few hundred nodes or less (so that a computation that would take a
-CPU-year can produce results in less than a day).
+specialized solutions are needed~\citep{dean2008mapreduce,zaharia2010spark}.
+There is a range of medium-sized problems that can be successfully tackled on a
+computer cluster with a small number of nodes or even taking advantage of a
+single multicore machine.
 
 The typical ad hoc solution to this problem is to save intermediate files on
-disk. Often the design of the computation itself takes several iterations as
-intermediate steps are improved. Thus, some intermediate results need to be
-recomputed. This involves a great amount of human management of the state of
-the computation, breaking it up into pieces, and issuing jobs on the cluster.
+disk. A limitation of this approach is that often the design of the computation
+itself takes several iterations as intermediate steps are improved. Thus, some
+intermediate results need to be recomputed. This involves a large amount of
+human management of the state of the computation, breaking it up into pieces,
+and issuing jobs on the cluster.
 
@@
 
-However, unlike make, jug is written in Python and tasks can consist of any
-Python function and its arguments. This can include running external commands
-and calling routines written in other languages as Python has many tools to
-interface with the rest of the system.
+However, unlike make, jug is written in Python, a general purpose programming
+language which is widely used in scientific programming \citep{prabhu2011a}.  A
+Task in Jug can consist of any Python function and its arguments. The task can
+include running external commands and calling routines written in other
+languages as Python has many tools to interface with the rest of the system
+\citep{beazley1996swig,beazley2003automated,behnel2011cython}.
 
-A jugfile (the file which specifies which tasks need to be run---named by
-analogy to the makefiles of make) is a simple Python script with some added
+A Jugfile (the file which specifies which tasks need to be run---named by
+analogy to the Makefiles of make) is a simple Python script with some added
 notations (we will see the only a few, small, changes are needed to transform a
@@
 module with zlib compression. As a special case, \code{numpy} arrays
-\citep{numpy} are saved to disk directly. This special case was introduced as
-\code{numpy} arrays are a very common data type in scientific programming and
-saving them directly allows for very fast saving and loading (they are
-represented on disk as a header followed by the binary information they
-contain).
+\citep{walt2011the} are saved to disk directly. This special case was
+introduced as \code{numpy} arrays are a very common data type in scientific
+programming and saving them directly allows for very fast saving and loading
+(they are represented on disk as a header followed by the binary information
+they contain).
 
@@
 The output of computation can be obtained from jug in three ways: (1) One can
-simply write a task that writes the output to a file in the desired format. (2)
+write a task that writes the output to a file in the desired format. (2)
 Outputs can be inspected interactively using the \emph{shell} subcommand. It is
@@
 (3) jug can be loaded as a library from a Python script and jug computation
-outputs can serve as inputs for further computation.
+outputs can serve as inputs for further computation. This can be performed from
+inside a Jupyter notebook \citep{kluyver2016jupyter}.
 
-The shell subcommand, invokes an IPython instance with all the objects in the
-jugfile loaded. IPython is an enhanced interactive shell for Python
-\citep{Perez2007}. A few functions are added to the namespace, in particular,
-\code{value} can load the results of a task object if necessary.
+The shell subcommand, invokes an Jupyter instance with all the objects in the
+jugfile loaded. The Jupyter console is an enhanced interactive shell for Python
+\citep{perez2007ipython}. A few functions are added to the namespace, in
+particular, \code{value} can load the results of a task object if necessary.
 
@@
 
-=========
-Jug Shell
-=========
-
 Available jug functions:
@@
     - load_all() : loads all objects
-
-Enjoy...
 
@@
 Figure~\ref{fig:jug-shell-interaction} shows a possible interaction session
-with jug shell. While having to explicitly load all the results is a bit
-bothersome, it can be better than pre-loading all of them. It is much faster at
-startup. Consider too that the user might not load more than a few objects. In
-some cases, loading all of the objects simultaneously might even be impossible
-due to memory constraints. Further, this allows exploration of the task
-structure for debugging.
-
-This is a feature which can be used to debug one's own code or to explore the
-results of a computation by others.
+with jug shell. While having to explicitly load all the results may be
+bothersome, it is both much faster at start up and the user might not load more
+than a few objects throughout their session. In some cases, loading all of the
+objects simultaneously might even be impossible due to memory constraints.
+Furthermore, this allows exploration of the task structure for debugging.
 
@@
 
-If a developer improves an intermediate step in a pipeline (e.g., fixes a bug)
+If a researcher improves an intermediate step in a pipeline (e.g., fixes a bug)
 and wishes to obtain new results of the computation, then all outputs from that
@@
 Therefore, jug adds support for result invalidation. When a results from a task
-are invalidated, all tasks which, directly or indirectly, depend on them are
-also marked for deletion.
+are invalidated, all tasks which (directly or indirectly) depend on them are
+also invalidated.
 
@@
 was done in a dynamic way, it could be hard to discover all dependencies (cf.\
-sumatra~\footnote{Sumatra is a tool to collect all dependencies that are
-necessary to run a particular computation. Its webpage is located at
-\url{http://neuralensemble.org/trac/sumatra}.}). Additionally, even minute
+sumatra, which does capture all dependencies that are necessary to run a
+particular computation \citep{davison2012automated}). Additionally, even minute
 refactoring of the code would lead to over eager recomputation. This could make
@@
 
-\section{Methods}
 \section{Example}
-
-I will present an example from my research. Although it introduces some
-unnecessary complexity, it has the advantage of being a real-life example of
-the usage of jug.
-
-The goal of this project was to demonstrate that using local features, in
-particular speeded-up robust features (henceforth,
-\textsc{surf})~\cite{surf-paper} improves over image-global features for
-classification of microscope images. A variation of standard \textsc{surf},
-\textsc{surf}-ref, was introduced and tested.
+This section presents an edited version of the code used in a previously
+published study of computer vision techniques for bioimage
+analysis~\cite{coelho2013determining}. The goal of this project was to
+demonstrate that using local features, in particular speeded-up robust features
+(henceforth, \textsc{surf}) improves over image-global features for
+classification of microscope images.
 
@@
 final result is the average of all ten values). This is known as
-cross-validation. In this case, I use another library (which I also wrote)
-which has a built-in method for jug-aware cross-validation.
+cross-validation.
 
@@
 from jug import TaskGenerator, CachedFunction
+import mahotas as mh
+from sklearn.cross_validation import KFold
 
-# pyslic is used for subcellular location analysis
-import pyslic
-
-# milk (machine learning toolkit)
-from milk.ext.jugparallel import nfoldcrossvalidation
-
-datasets = [ 'data0', 'data1', ]
-featuresets = [ 'field-dna+', 'surf', 'surf-ref', ]
-
-def load(dset):
+def load():
     '''
     This function assumes that the images are stored in
-    ../data/dset with filenames matching the pattern
-    label-(protein|dna)-([0-9]).tiff
+    data/<dset> with filenames matching the pattern
+    label-protein-([0-9]).tiff
     '''
@@
     images = []
-    base = '../data/'+dset
+    base = './data/'
     for path in listdir(base):
@@
         # This save memory.
-        im = pyslic.Image({
-                'protein': base + path,
-                'dna' : base + path.replace('protein','dna')
-                })
-        im.label = label
+        im = (base + path, label)
         images.append(im)
@@
 
-# pyslic.computefeatures takes an image and an identifier of the
-# feature set and returns a set of features (as a numeric array).
-#
-# TaskGenerator can be used directly to adapt existing functions
-computefeatures = TaskGenerator(pyslic.computefeatures)
+# TaskGenerator can be used with any function, including builtins:
+jug_sum = TaskGenerator(sum)
 
-# or it can be used as a decorator on a newly defined function
+@TaskGenerator
+def computefeatures(im):
+    im = mh.imread(im)
+    return mh.features.haralick(im)
+
 @TaskGenerator
 def output(cmatrix, dset, fset):
-    output = file('results/cmatrix-'+dset+'-'+fset+'.txt', 'w')
-    print >>output, cmatrix
-    output.close()
+    with open('results/cmatrix-'+dset+'-'+fset+'.txt', 'w') as output:
+        output.write(cmatrix)
+        output.write('\n')
 
+@TaskGenerator
+def train_test1(features, labels, train, test):
+    from sklearn import linear_model
+    from sklearn import metrics
+    clf = linear_model.LogisticRegression()
+    clf.fit(features[train], labels[train])
+    preds = clf.predict(features[test])
+    return metrics.confusion_matrix(labels[test], preds)
+
+images = CachedFunction(load)
 features = []
-for dset in datasets:
-    images = CachedFunction(load, dset)
-    labels = [im.label for im in images]
-    for fset in featuresets:
-        features = [computefeatures(im, fset) for im in images]
+images = []
+for im,label in images:
+    labels.append(label)
+    features.append(haralick(im))
+cmats = []
+for train,test in KFold(len(features), n_folds=10):
+    cmats.append(
+        train_test1(features, labels, train, test))
+return jug_sum(cmats)
 cmatrix = nfoldcrossvalidation(features, labels)
@@
 \url{https://www.github.com/luispedro/jug}.} An open mailing-list provides
-support and reported bugs are generally fixed within hours.
+support and reported bugs are generally fixed within hours or days.
 
-Jug includes a full test suite. There are no known bugs.
+Jug includes a full test suite with continuous integration, which are best
+practices \citep{wilson2014best} which can improve productivity
+\citep{vasilescu2015quality}. There are no known bugs.
 
@@
 
-\citet{Fomel2007} proposed a system built on top of Scons which shared
-superficially similar design. Scons, like Make, is a build system. It supports
-spawning parallel jobs using multiple threads. In the original use case, the
-tasks are delegated to other commands and the operating system can take care of
-parallelism. If, however, the tasks are computationally intensive in Python,
-contention for the global interpreter lock will limit the amount of real
-parallelism.\footnote{In the most commonly used Python interpreters, there is a
-lock that prevents more than one thread from simultaneously executing
-interpreted code, although they can execute non-interpreted code, as calling an
-external programme or certain external libraries.} Ruffus is another
-Python-based solution which supports parallel execution of pipelines
-\citep{Goodstadt16092010}. There is a strong dependency on the filesystem,
-which makes its use very different from jug usage, which strives to make its
-usage similar to straight Python usage. Using the Ruby programming language,
-\citet{mishima2011} proposed Pwrake, which supports parallel execution of
-bioinformatics workflows written (or accessible) in that language.
+Several pipeline tools have been used (for a recent review, see the work of
+\citet{leipzig2016a}). \citet{2007reproducible} proposed a system built on
+top of Scons which shared superficially similar design. Scons, like Make, is a
+build system. It supports spawning parallel jobs using multiple threads. In the
+original use case, the tasks are delegated to other commands and the operating
+system can take care of parallelism. If, however, the tasks are computationally
+intensive in Python, contention for the global interpreter lock will limit the
+amount of real parallelism.\footnote{In the most commonly used Python
+interpreters, there is a lock that prevents more than one thread from
+simultaneously executing interpreted code, although they can execute
+non-interpreted code, as calling an external programme or certain external
+libraries.} Ruffus is another Python-based solution which supports parallel
+execution of pipelines \citep{goodstadt2010ruffus}. Using the Ruby programming
+language, \citet{mishima2011} proposed Pwrake, which supports parallel
+execution of bioinformatics workflows written (or accessible) in that language.
+Snakemake~\citep{kster2012snakemakea} also improves over Make by providing more
+complex rules and automatic interaction with a high performance compute
+cluster. \citet{sadedin2012bpipe} proposed Bpipe, a tool based on a
+domain-specific language around shell scripting. One large difference of Jug
+compared to these tools is that Jug is tightly integrated with Python code.
+Thus, while it lacks support for directly spawning processes like Make, it is
+easier to call Python-based functions.
 
-Of interest, is also the \textsc{IncPy} implementation of a Python interpreter
-with automatic persistent memoization~\cite{Guo_IncPy,Guo_IncPy_pre}. The
-arguments made for that system apply to Jug as well, with a few differences.
-Their system implements automatic memoization, while, using Jug, the user needs
-to automatically annotate functions to memoize using \texttt{TaskGenerator}.
-This extra control can be necessary to avoid a proliferation of intermediate
-results when these are very large (often a very large intermediate output is
-not necessary, and only a summary must be kept). Additionally, Jug supports
-running tasks in parallel, a functionality that is absent from \textsc{IncPy}.
+Also relevant is the \textsc{IncPy} implementation of a Python interpreter with
+automatic persistent memoization~\cite{guo2010towards,guo2011using}
+(unfortunately, no longer maintained). The advantages of that system apply to
+Jug as well, with a few differences. Their system implements automatic
+memoization, while, using Jug, the user needs to automatically annotate
+functions to memoize using \texttt{TaskGenerator}.  This extra control can be
+necessary to avoid a proliferation of intermediate results when these are very
+large (often a very large intermediate output is not necessary, and only a
+summary must be kept) at the cost of increased overhead for the programmer.
+Additionally, Jug supports running tasks in parallel, a functionality that is
+absent from \textsc{IncPy}.
+
+Finally, joblib and memo \citep{moreno2016improving} use mechanisms similar to
+Jug for in-process and in-memory memoization and parallelization.
 
@@
 
-Unlike most previous tools for reproducible research, jug focuses on the
-development of the computation as much as its communication. In fact, when it
-comes to communication other tools might be better suited as they combine
-written exposition with computer code. They can be used in combination with jug
-as they do not often provide caching and distribution, needed for large
-projects. While the simplest mode of jug use is to have a single
-\textit{jugfile}, a jugfile can import the results of another. For example, a
-first jugfile might perform all heavy-duty computation and be run on a
-computing cluster. A second script could be embedded inside an executable paper
-to generate tables and plots \citep{Delescluse2011}. This faster script could
+Jug focuses on the development of the computation as much as its communication.
+In fact, when it comes to communication other tools might be better suited as
+they combine written exposition with computer code. They can be used in
+combination with jug as they do not often provide caching and distribution,
+needed for large projects. While the simplest mode of jug use is to have a
+single \textit{jugfile}, a jugfile can import the results of another. For
+example, a first jugfile might perform all heavy-duty computation and be run on
+a computing cluster. A second script could be embedded inside an executable
+paper to generate tables and plots \citep{2011making}. This faster script could
 be run as part of building the paper (while the longer script would take too
@@
 Jug improves the pipeline development experience and makes the programming
-researcher more productive and less error-prone. As it is based on Python, a
-widely used scientific programming language, it can be used in a known
-environment without a long learning process.
+researcher more productive and less error-prone. As many scientists now spend a
+significant fraction of their time performing computational work
+\citep{prabhu2011a,hannay2009how}, increasing their productivity in this task
+can have significant effects.  Jug is based on Python, a widely used scientific
+programming language, it can be used in a known environment without a high
+learning curve.
 
@@
 have much higher costs in terms of development time (code must be written
-especially for the framework) and overhead. For scientists using small to
-medium-sized cluster (hundreds, not thousands, of nodes), with a shared
-file-system (as opposed to the distributed file-systems necessary for truly
-large amounts of data) jug provides a better trade-off than those
-higher-powered alternatives in a familiar programming language.
+especially for the framework) and overhead. For scientists who want to quickly
+adapt pre-existing code to run on a cluster or even a multi-core machine, jug
+provides a better trade-off than those higher-powered alternatives in a
+familiar programming language.
 
